State of the Practice for Medical Imaging Software

### 2022-07-18 StateOfPractice/Papers/MedImageSoft_SOP-arXiv/MedImageSoft_SOP_AcademicRadiology.tex
Edits to pain points section, specifically P1 and P2

Based on interviews with nine developers (described in
Section~\ref{sec_interview_methods}), we answer three research questions:
\rqref{RQ_PainPoints}) What are the pain points for developers working on
research software projects?; \rqref{RQ_ComparePainPoints}) How do the pain
points of developers from MI compare to the pain points for research software in
general?; and \rqref{RQ_Concerns}) For MI developers what specific best
practices are taken to address the pain points and software quality concerns?
The identified pain points relate to lack of time and funding, technology
hurdles, improving correctness, and improving usability.  We go through each
pain points and include citations that contrast the MI experience with
observations from other domains.  We also cover potential ways to address the
pain points, as promoted by the community in this section, and based on our
experience in Section~\ref{ch_recommendations}.  In this section we also discuss
some MI developer strategies for improving maintainability and reproducibility.
These two qualities were not explicitly idenfied as pain points, but ways to
improve them were still part of our interview discussions. The MI interviewee
practices to address pain points and improve quality can potentially be emulated
by other the MI software packages.  Moreover, these practices may provide
examples that can be followed by other research software domains.

\citet{PintoEtAl2018} lists some pain points that did not come up in our
conversations with LBM developers [CHECK  THESE]: Cross-platform compatibility, interruptions
while coding, scope bloat, lack of user feedback, hard to collaborate on
software projects, and aloneness. \citet{WieseEtAl2019} repeat some of the
previous pain points and add the following: dependency management, data handling
concerns (like data quality, data management and data privacy), reproducibility,
and software scope determination. Although LBM developers did not mention these
pain points, we cannot conclude that they are not relevant for LBM software
development, since we only interviewed four LBM developers for about an hour each.

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
Development Time:} A developer of pyLBM noted that their small development team
has a lack of time to implement new features. Small development teams are common
for LBM software packages (as shown in the measurement table excerpt in
Figure~\ref{measurement_template_image}). Other domains of research software
also experience the lack of time pain point \citep{PintoEtAl2018, PintoEtAl2016,
WieseEtAl2019}. Developers identified a resource related pain point caused by a
lack of funding and time.

@@

\begin{itemize}
\item Shifting from development to maintenance when the team does not have
enough developers for building new features and fixing bugs at the same time;
\item Improving documentation to save time answering users' and developers'
questions;
\item Supporting third-party plugins and extensions; and,
\item Using GitHub Actions for CI/CD to save time.
\end{itemize}

Many interviewees thought lack of fundings and lack of time were their most
significant obstacles. The interviewees from \textit{3D Slicer} team and
\textit{OHIF} team pointed out that it was more challenging to get fundings for
software maintenance as opposed to research. The interviewee from the
\textit{ITK-SNAP} team thought more fundings was a way to solve the lack of time
problem, because they could hire more dedicated developers. On the other hand,
the interviewee from the \textit{Weasis} team did not feel that fundings could
solve the same problem, since he still would need a lot of time to supervise the
project.  No interviewee suggested any solution to bring extra funding to the
project. However, they provided ideas to save time, such as better
documentation, third-party plugins, and good CI/CD tools.

\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
Incentive and Funding:} The TCLB developer noted a lack of incentives and
funding in academia for developing widely used research software. This problem
has also been noted by others \citep{gewaltig2012quality, Goble2014,
KaterbowAndFeulner2018}.  \citet{WieseEtAl2019} reported developer pains related
to publicity, since publishing norms have historically made it difficult to get
credit for creating software.  As studied by \citet{HowisonAndBullard2016},
research software (specifically biology software, but the trend likely applies
to other research software domains) is infrequently cited. \citet{PintoEtAl2018}
also mentions the lack of formal reward system for research software.

\begin{itemize}
\item Licensing the software to commercial companies that integrate it into
their products;
\end{itemize}

@@

@@

- add in ACs, borrow from previous paper.

>>>>

Based on interviews with nine developers (described in
Section~\ref{sec_interview_methods}), we answer three research questions (first
mentioned in Section~\ref{sec_motivation}): \rqref{RQ_PainPoints}) What are the
pain points for developers working on research software projects?;
\rqref{RQ_ComparePainPoints}) How do the pain points of developers from MI
compare to the pain points for research software in general?; and
\rqref{RQ_Concerns}) For MI developers what specific best practices are taken to
address the pain points and software quality concerns?

Our interviews identify pain points that relate to lack of time and funding,
technology hurdles, improving correctness, and improving usability.  In this
section, we go through each pain point and contrast the MI experience with
observations from other domains.  We also cover potential ways to address the
pain points, as promoted by the community.  (Later, in
Section~\ref{ch_recommendations}, we propose additional pain mitigation
strategies based on our experience.)  In addition to pain points, we summarize
MI developer strategies for improving maintainability and reproducibility.
Although these two qualities were not explicitly idenfied as pain points, they
were discussed during our interviews.  The interviewee's practices for
addressing pain points and improving quality can potentially be emulated by
other MI developers. Moreover, these practices may provide examples that can be
followed by other research software domains.

\citet{PintoEtAl2018} lists some pain points that did not come up in our
conversations with MI developers [CHECK  THESE, after writing all the PPs]:
Cross-platform compatibility, interruptions while coding, scope bloat, lack of
user feedback, hard to collaborate on software projects, and aloneness.
\citet{WieseEtAl2019} repeat some previous pain points and add the following:
dependency management, data handling concerns (like data quality, data
management and data privacy) [Actually mentioned by MI software],
reproducibility, and software scope determination. Although LBM developers did
not mention these pain points, we cannot conclude that they are not relevant for
LBM software development, since we only interviewed four LBM developers for
about an hour each.

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
Development Time:} Many interviewees thought lack of time and lack of funding
(discussed next) were their most significant obstacles. Other domains of
research software also experience the lack of time pain point
\citep{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}. Our study of LBM software
\citep{SmithEtAl2022} also highlighted lack of time as a frequently mentioned
pain point.

@@

\begin{itemize}
\item Shifting from development to maintenance when the team does not have
enough developers for building new features and fixing bugs at the same time;
\item Improving documentation to save time answering users' and developers'
questions;
\item Supporting third-party plugins and extensions; and,
\item Using GitHub Actions for continunous integration and continuous
development.
\end{itemize}

\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
Incentive and Funding:} As mentioned above, developers felt the pain of having
to attract funding the development and maintenance of their software projects.
For instance, the interviewees from \textit{3D Slicer} and \textit{OHIF} pointed
out that it was more challenging to get fundings for software maintenance as
opposed to research. The interviewee from the \textit{ITK-SNAP} team thought
more fundings was a way to solve the lack of time problem, because they could
hire more dedicated developers. On the other hand, the interviewee from the
\textit{Weasis} team did not feel that fundings could solve the same problem,
since he still would need a lot of time to supervise the project.

Funding challenges have also been noted by others \citep{GewaltigAndCannon2012,
Goble2014, KaterbowAndFeulner2018, SmithEtAl2022}. \citet{WieseEtAl2019}
reported developer pains related to publicity, since publishing norms have
historically made it difficult to get credit for creating software.  As studied
by \citet{HowisonAndBullard2016}, research software (specifically biology
software, but the trend likely applies to other research software domains) is
infrequently cited. \citet{PintoEtAl2018} also mentions the lack of formal
reward system for research software.

An interviewee proposed an idea for increasing funding: Licensing the software
to commercial companies that integrate it into their products.

@@

%\citet{WieseEtAl2019} mention data handling concerns (like data quality, data management and data privacy)

@@

- add in ACs, borrow from previous paper.
- mention the solutions proposed by LBM developers, like design for change
